# 2B Lab 1 Week 2  {.unnumbered}


```{r include=FALSE}
library(webexercises)
```

**This is the pair coding activity related to @sec-paired.**



## Task 1: Open the R project for the lab {.unnumbered}

## Task 2: Create a new `.Rmd` file {.unnumbered}

... and name it something useful. If you need help, have a look at @sec-rmd.

## Task 3: Load in the library and read in the data {.unnumbered}

The data should already be in your project folder. If you want a fresh copy, you can download the data again here: [data_pair_coding](data/data_pair_coding.zip "download").

We are using the packages `rstatix`, `tidyverse`, `qqplotr`, `lsr` today.
Make sure to load `rstatix` in before `tidyverse`.

We also need to read in `dog_data_clean_wide.csv`. Again, I've named my data object `dog_data_wide` to shorten the name but feel free to use whatever object name sounds intuitive to you. 

For the plot, we will need the data in long format. We can either read in `dog_data_clean_long.csv` to take a shortcut, or wrangle the data from `dog_data_wide`. I've taken the shortcut and named my data object `dog_data_long`.



```{r reading in data for me, echo=FALSE, message=FALSE}
library(rstatix)
library(tidyverse)
library(lsr)

dog_data_wide <- read_csv("data/dog_data_clean_wide.csv")
dog_data_long <- read_csv("data/dog_data_clean_long.csv")
```

## Task 4: Tidy data for a paired t-test {.unnumbered}

Not much tidying to do for today. 

Pick a variable of interest and select the pre- and post-scores, and calculate the difference score. Store them in a separate data object with a meaningful name.

I will use Loneliness as an example and call my data object `dog_lonely`. Regardless of your chosen variable, your data object should look like/ similar to the table below. 

```{r echo=FALSE, warning=FALSE, message=FALSE}
dog_lonely <- dog_data_wide %>% 
  select(RID, Loneliness_pre, Loneliness_post) %>% 
  mutate(Loneliness_diff = Loneliness_post - Loneliness_pre)

head(dog_lonely, n = 5)
```

In `dog_data_long`, we want to turn `Stage` into a factor so we can re-order the labels (i.e., "pre" before "post").


::: {.callout-caution collapse="true" icon="false"}

## Solution for Tasks 3 and 4

```{r}
## Task 3
library(rstatix)
library(tidyverse)
library(lsr)

dog_data_wide <- read_csv("data/dog_data_clean_wide.csv")
dog_data_long <- read_csv("data/dog_data_clean_long.csv")

## Task 4
dog_lonely <- dog_data_wide %>% 
  select(RID, Loneliness_pre, Loneliness_post) %>% 
  mutate(Loneliness_diff = Loneliness_post - Loneliness_pre)

dog_data_long <- dog_data_long %>% 
  mutate(Stage = factor(Stage,
                        levels = c("pre", "post")))
```


:::

## Task 5: Compute descriptives {.unnumbered}

We want to determine the mean and sd of:

* the pre-scores
* the post-scores, and
* the difference scores

Store them in a data object called `descriptives`.

::: {.callout-caution collapse="true" icon="false"}

## Solution

```{r}
descriptives <- dog_lonely %>% 
  summarise(mean_pre = mean(Loneliness_pre),
            sd_pre = sd(Loneliness_pre),
            mean_post = mean(Loneliness_post),
            sd_post = sd(Loneliness_post),
            diff = mean(Loneliness_diff),
            sd_diff = sd(Loneliness_diff))
```

:::


## Task 6: Check assumptions {.unnumbered}


#### Assumption 1: Continuous DV {.unnumbered} 

Is the dependent variable (DV) continuous? Answer: `r longmcq(sample(c(answer = "Yes. The DV is the difference in loneliness scores and it is continuous", x = "No. The DV is the difference in loneliness scores but it is categorical", x = "No. The DV is the pre- and post-stages and it is categorical")))`


#### Assumption 2: Data are independent {.unnumbered} 

Each pair of values in the dataset has to be independent, meaning each pair of values needs to be from a separate participant. Answer: `r mcq(c(answer = "yes", x = "no"))`

#### Assumption 3: Normality {.unnumbered}

Looking at the violin-boxplots below, do you think the assumption of normality holds?

::: {.callout-note}
The axis label of Plot 2 turned out to be quite long here. I've used the escape character `\n` to break it up across 2 lines.
:::

```{r eval = FALSE, message = FALSE}
## Plot 1
ggplot(dog_data_long, aes(x = Stage, y = Loneliness, fill = Stage)) +
  geom_violin(alpha = 0.5) +
  geom_boxplot(width = 0.4, alpha = 0.8) +
  scale_fill_viridis_d(guide = "none") +
  theme_classic() +
  labs(x = "Time point", y = "mean Loneliness Scores")

## Plot 2
ggplot(dog_lonely, aes(x = "", y = Loneliness_diff)) +
  geom_violin(fill = "#21908C", alpha = 0.5) +
  geom_boxplot(fill = "#21908C", width = 0.4) +
  theme_classic() +
  labs(x = "",
       y = "Difference in mean Loneliness scores \nbetween pre- and post- intervention") # \n forces a manual line break in the axis label

```

```{r fig.cap="Plots displayed to assess normality assumption", echo=FALSE}
library(patchwork)

## Plot 1
p1 <- ggplot(dog_data_long, aes(x = Stage, y = Loneliness, fill = Stage)) +
  geom_violin(alpha = 0.5) +
  geom_boxplot(width = 0.4, alpha = 0.8) +
  scale_fill_viridis_d(guide = "none") +
  theme_classic() +
  labs(x = "Time point", y = "mean Loneliness Scores") +
  ggtitle("Plot 1")

## Plot 2
p2 <- ggplot(dog_lonely, aes(x = "", y = Loneliness_diff)) +
  geom_violin(fill = "#21908C", alpha = 0.5) +
  geom_boxplot(fill = "#21908C", width = 0.4) +
  theme_classic() +
  labs(x = "",
       y = "Difference in mean Loneliness scores \nbetween pre- and post- intervention") +
  ggtitle("Plot 2")

p1 + p2 + plot_layout(nrow = 1)
```




Answer:

`r longmcq(c(x = "yes, because both pre- and post-scores in Plot 1 are approximately normally distributed", answer = "yes, because the difference scores in Plot 2 are approximately normally distributed", x = "no, because both pre- and post-scores in Plot 1 are extremely skewed", x = "no, because the difference scores in Plot 2 are extremely skewed"))`

#### Conclusion from assumption tests {.unnumbered}

With all assumptions tested, which statistical test would you recommend for this analysis?

Answer: `r longmcq(c(answer = "All assumptions held. We will conduct a paired-samples t-test.", x = "The assumption of normality was violated. We will conduct a Wilcoxon signed-rank test."))`



## Task 7: Computing a paired-sample t-test with effect size & interpret the output {.unnumbered}


* **Step 1**: Compute the paired-sample t-test. The structure of the function is as follows:

```{r eval=FALSE}
t.test(your_data$var1, your_data$var2, paired = TRUE)
```


::: {.callout-caution collapse="true" icon="false"}

## Solution

```{r eval=FALSE}
t.test(dog_lonely$Loneliness_pre, dog_lonely$Loneliness_post, paired = TRUE)
```

:::


* **Step 2**: Calculate an effect size

Calculate Cohen's D. The structure of the function is as follows:

```{r eval=FALSE}
cohensD(your_data$var1, your_data$var2, method = "paired")
```


::: {.callout-caution collapse="true" icon="false"}

## Solution

```{r eval=FALSE}
cohensD(dog_lonely$Loneliness_pre, dog_lonely$Loneliness_post, method = "paired")
```

:::


* **Step 3**: Interpreting the output

Below are the outputs for the **descriptive statistics** (table), **paired-samples t-test** (main output), and **Cohen’s D** (last line starting with [1]). Based on these, write up the results in APA style and provide an interpretation.

```{r echo=FALSE}
descriptives

t.test(dog_lonely$Loneliness_pre, dog_lonely$Loneliness_post, paired = TRUE)

cohensD(dog_lonely$Loneliness_pre, dog_lonely$Loneliness_post, method = "paired")
```

We hypothesised that there would be a significant difference between Loneliness measured before (*M* = `r fitb("2.04")`, *SD* = `r fitb("0.53")`) and after (*M* = `r fitb("1.91")`, *SD* = `r fitb("0.53")`) the dog intervention. On average, participants felt less lonely after the intervention (*M~diff~* = `r fitb("0.13")`, *SD~diff~* = `r fitb("0.23")`). 
Using a paired-samples t-test, the effect was found to be `r mcq(c(answer = "significant", x = "non-significant"))` and of a `r mcq(c(answer = "small", x = "medium", x = "large"))` magnitude, *t*(`r fitb("283")`) = `r fitb("9.26")`, *p* `r fitb("< .001")`, *d* = `r fitb("0.55")`. We therefore `r mcq(c(x = "fail to reject the null hypothesis", answer = "reject the null hypothesis in favour of H1"))`.

